---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)
```

# rkaf

`rkaf` is an R package for **Kolmogorov-Arnold Fourier Networks** using
the [`torch`](https://torch.mlverse.org/) backend.

The package provides a modern R interface for KAF models, including:

- regression
- binary classification
- multiclass classification
- formula and matrix interfaces
- mini-batch training
- validation splits
- early stopping
- automatic standardization
- best-model restoration
- KAF-specific diagnostics

The goal of `rkaf` is to make KAF-style neural networks accessible to R users
without requiring Python, `reticulate`, or custom training loops.

## Installation

You can install the development version locally with:

```r
devtools::install()
```

Or from a local package directory:

```r
devtools::load_all()
```

`rkaf` depends on `torch`. If `torch` is not already configured, run:

```r
torch::install_torch()
```

## Basic regression example

```{r regression-example}
library(rkaf)

set.seed(123)
torch::torch_manual_seed(123)

x <- as.matrix(seq(-1, 1, length.out = 128))

y <- sin(8 * pi * x) +
  0.35 * cos(3 * pi * x) +
  0.15 * x^2

fit <- kaf_fit(
  x = x,
  y = y,
  hidden = c(32, 32),
  num_grids = 8,
  use_layernorm = FALSE,
  epochs = 100,
  lr = 1e-3,
  verbose = FALSE,
  seed = 123
)

fit
```

```{r regression-prediction-plot}
pred <- predict(fit, x)

plot(
  x,
  y,
  type = "l",
  lwd = 2,
  xlab = "x",
  ylab = "f(x)",
  main = "KAF regression example"
)

lines(x, pred, lwd = 2, lty = 2)

legend(
  "topright",
  legend = c("Observed", "Predicted"),
  lty = c(1, 2),
  lwd = 2,
  bty = "n"
)
```

## Formula interface

`rkaf` also supports a formula interface for tabular data.

```{r formula-regression}
fit_mtcars <- kaf_fit_formula(
  mpg ~ wt + hp + cyl,
  data = mtcars,
  hidden = c(16, 16),
  num_grids = 8,
  epochs = 50,
  verbose = FALSE,
  seed = 123
)

fit_mtcars
```

```{r formula-regression-predict}
head(predict(fit_mtcars, mtcars))
```

## Binary classification

For factor, character, or logical targets with two classes, `rkaf` can fit a
binary classifier automatically.

```{r binary-classification}
df <- mtcars
df$high_mpg <- factor(
  ifelse(df$mpg > median(df$mpg), "yes", "no"),
  levels = c("no", "yes")
)

fit_binary <- kaf_fit_formula(
  high_mpg ~ wt + hp + cyl,
  data = df,
  hidden = c(16, 16),
  num_grids = 8,
  epochs = 50,
  verbose = FALSE,
  seed = 123
)

fit_binary
```

Predicted probabilities:

```{r binary-prob}
head(predict(fit_binary, df, type = "prob"))
```

Predicted classes:

```{r binary-class}
head(predict(fit_binary, df, type = "class"))
```

## Multiclass classification

For factor targets with more than two classes, `rkaf` fits a multiclass model.

```{r multiclass-classification}
fit_iris <- kaf_fit_formula(
  Species ~ .,
  data = iris,
  hidden = c(16, 16),
  num_grids = 8,
  epochs = 50,
  verbose = FALSE,
  seed = 123
)

fit_iris
```

Class probabilities:

```{r multiclass-prob}
head(predict(fit_iris, iris, type = "prob"))
```

Predicted classes:

```{r multiclass-class}
head(predict(fit_iris, iris, type = "class"))
```

## Validation and early stopping

`kaf_fit()` supports validation splits, mini-batching, and early stopping.

```{r validation-example}
fit_val <- kaf_fit(
  x = x,
  y = y,
  hidden = c(32, 32),
  num_grids = 8,
  use_layernorm = FALSE,
  epochs = 300,
  batch_size = 32,
  validation_split = 0.2,
  patience = 30,
  verbose = FALSE,
  seed = 123
)

fit_val
```

```{r validation-plot}
plot(fit_val)
```

## KAF diagnostics

You can inspect the learned balance between the base/GELU branch and the
Fourier branch.

```{r diagnostics-scales}
scales <- extract_kaf_scales(fit)
head(scales)
```

```{r diagnostics-fourier}
fourier_params <- extract_fourier_params(fit, layer = 1)
head(fourier_params)
```

```{r diagnostics-plot}
plot_kaf_scales(fit, layer = 1, type = "ratio")
```

## Low-level torch interface

Advanced users can directly create KAF torch modules.

```{r low-level-interface}
model <- nn_kaf(
  layers = c(4, 16, 16, 1),
  num_grids = 8
)

x_tensor <- torch::torch_randn(10, 4)
y_tensor <- model(x_tensor)

y_tensor$shape
```

## Current package status

The package currently includes:

```r
devtools::test()
devtools::check()
```

with full test coverage for the core API, including regression, binary
classification, multiclass classification, formula handling, prediction,
training utilities, and diagnostics.

## References and attribution

`rkaf` is an independent R implementation of Kolmogorov-Arnold Fourier
Networks using the `torch` backend.

The package is based on the KAF architecture proposed in:

> Zhang, J., Fan, Y., Cai, K., & Wang, K. (2025).  
> Kolmogorov-Arnold Fourier Networks.  
> arXiv:2502.06018.

The implementation was also informed by the public Python reference
implementations:

- `kolmogorovArnoldFourierNetwork/KAF`
- `kolmogorovArnoldFourierNetwork/kaf_act`

No Python dependency is required by `rkaf`.

## License

MIT.
